---
title: "Untitled"
author: "Felipe Rodriguez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set the working directory to the root of your DSC 520 directory
## Load the `data/r4ds/heights.csv` to
```{r}
setwd("/Users/feliperodriguez/Library/CloudStorage/OneDrive-BellevueUniversity/Github/dsc520")
heights_df <- read.csv("data/r4ds/heights.csv")
```

## Load the ggplot2 library
```{r}
library(ggplot2)
library(dplyr)
```

## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome

```{r}
age_lm <- lm(earn ~ age, data=heights_df)
```

## View the summary of your model using `summary()`
```{r}
summary(age_lm)
```

## Creating predictions using `predict()`
```{r}
age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age=heights_df$age)
```


## Plot the predictions against the original data
```{r}
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y= earn, x= age))
```

```{r}
mean_earn <- mean(heights_df$earn)
```

## Corrected Sum of Squares Total
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
```

## Corrected Sum of Squares for Model
```{r}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
```

## Residuals
```{r}
residuals <- heights_df$earn - age_predict_df$earn
```

## Sum of Squares for Error
```{r}
sse <- sum(residuals^2)
```

## R Squared R^2 = SSM\SST
```{r}
r_squared <- ssm/sst
```

## Number of observations
```{r}
n <- nrow(heights_df)
```

## Number of regression parameters
```{r}
p <- 2
```

## Corrected Degrees of Freedom for Model (p-1)
```{r}
dfm <- p-1
```

## Degrees of Freedom for Error (n-p)
```{r}
dfe <- n-p
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r}
dft <- n - 1
```

## Mean of Squares for Model:   MSM = SSM / DFM
```{r}
msm <- ssm/dfm
```

## Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse <- sse/dfe
```

## Mean of Squares Total:   MST = SST / DFT
```{r}
mst <- sst/dft
```

## F Statistic F = MSM/MSE
```{r}
f_score <- mse/mse
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}
adjusted_r_squared <- 1 - (1 - r_squared) * (n-1) / (n - p)
```

## Calculate the p-value from the F distribution
```{r}
p_value <- pf(f_score, dfm, dft, lower.tail=F)
```



